Demo for a Quality of Service networking with Software Defined Network.

### 2017-03-15 report/paper.tex

\section{Introduction}\label{sec:intro}
Software-defined networking (SDN) is a way to manage networks that separates the control from the forwarding plane.
SDN offers a centralized view of the network, giving an SDN Controller the ability to act as the “brain” of the network.
The SDN Controller relays on the information given by switches and routers via southbound APIs, and acts with what the applications set via northbound APIs.
One of the most well-known protocols used by SDN Controllers is OpenFlow, widely accepted even if it is not the only SDN implementation available. 
Centralized, programmable SDN environments can easily adjust to the rapidly changing business needs and can reduce costs, restrict
wasteful provisioning, as well as provide flexibility and innovation for networks.
%corretto
\subsection{OpenFlow}
  During 2006 ,thanks to M. Casado, a new network architecture was developed, called Ethane.
  On this conncept was born, during 2011, the Open Network Foundation, a no profit association that would innovate SDN and standardize OpenFlow protocol whit relatives tecnology \cite{ONF}.
  The first version of OpenFlow is 1.0.0 was released on 2009 but the last standardized version v1.5.1 was approved only on march 2015 \cite{ONF_report}.
  %corrects
  %todo spostare l'immagine correttamente
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.58]{images/of.jpg}
 \caption{\emph{OpenFlow}}
 \label{fig:topo}
\end{figure}
  \\
  \textbf{Openflow 1.0 messages}

@@

\subsection{Router Board}
\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.30]{images/rboard.png}
  \caption{\emph{Experience scenario}}
  \label{fig:topo}
\end{figure}

@@

When we set up this option \texttt{OFPP_FLOOD} like output port's action in a table's switch row if some spanning_tree algorithm is working
the \texttt{OFPP_FLOOD} is only theoretical in the sense that the algorithm excludes some ports during the flooding option, in order to avoid loop\cite{pox}.
\end{itemize}
\subsection{Real Lab Scenario}
The topological scenario in which we have set up experience is as follows:
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/topo.png}
 \caption{\emph{Experience scenario}}
 \label{fig:topo}
\end{figure}

@@

  \begin{figure}[!h]
 \centering
 \includegraphics[scale=0.30]{images/topology.png}
 \caption{\emph{Experience scenario}}
 \label{fig:topo}
  \end{figure}

\subsection{Traffic's Analysis}\label{subsec:traf}
We used laptop connected via
GigabitEthernet interface and Ethernet cable cat.5a to receive the traffic from the destination host.
Our laptop has been connected via Wi-Fi to our access point via protocol 802.11n.
No other traffic sources were active during the testing, so the router was handling only
the traffic from laptop when sniffing traffic. The connection used was an ADSL2+ via copper.
We did 3 type of traffic analysis: Gaming traffic, Skype® and torrent traffic. We analyzed
\#Mb for the gaming traffic, 70 Mb for the VoIP traffic and 3GB for the torrent traffic.
The analysis involved the average number of packets per second, the packet size, the protocol (TCP or UDP).

@@

\subsubsection{Delay}
Monitoring round-trip time provides important insights for network troubleshooting and traffic engineering.
The  common  monitoring  technique  is  to  actively  send  probe packets  from  selected  vantage  points  hosts  or  middleboxes.
In traditional  networks, the control over the network routing is
limited,  making  it  impossible  to  monitor  every  selected  path. The  emerging  concept  of  Software  Defined  Networking  simplifies
network  control.  However,  OpenFlow,  the  common  SDN protocol,   does   not   support   RTT   monitoring   as   part   of   its
specification. In this experience we have tryed to emply an tecnic to escape this problem\cite{rtt}.
\newline The idea is to measure the time that occur from sending a packet from controller to a switch and the time in which the packet return from
an other switch.
\\
\newline First we have to discovery round trip time from controller to every single switch on our topology.
In order to do this on the controller we start a timer, we send a proper packet to the switch specifying the \texttt{OFPP_CONTROLLER - Send to the controller}
option and when the message return to the controller we stop the timer and calculate RTT for the switch.
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/rtt2.png}
 \caption{\emph{RTT between controller and OF switch}}
 \label{fig:topo}
\end{figure}
\newline
If we need to calculate the delay on a specific link, the rtt has to be calculated with the headed switch of this link.
\\
\newline
Now we create a proper message with unique id. number, let's start a timer \texttt{TM1} and then send this message to a first switch specifying the output port.
Then when the packet comes to the second switch it identify the particular message and sends to the controller that stops it's dedicated timer \texttt{TM2}.
\newline This part required a particular flow entry on the second switch in order to send to the controller only some particular packet.
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/rtt1.png}
 \caption{\emph{Pck from controller to itself following a particular way.}}
 \label{fig:topo}
\end{figure}
\\
\newline At this point we can calculate the delay of the link.
\\
\\
 delay = (TM2 - TM1) - RTT1/2 - RTT2/2
\\
\\
Theoretically this method must works but in practice there are a lot of delay introduced from computational congestion on the virtualized controller.
What we observe is a estimated delay like a few seconds which is not the case, in fact if we made some \texttt{ping test} from one host to another host
in this SDN network we realize delay around 3ms.
\\
\\
The following picture shows the visual output that we can obtain from delay's graph. In this graph we storage only the core information;
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.26]{images/delay.png}
 \caption{\emph{Delay's graph.}}
 \label{fig:topo}
\end{figure}

		\subsubsection{Link Capacity}

>>>>

\section{Introduction}\label{sec:intro}
Software-defined networking (SDN) is a way to manage networks that separates the control from the forwarding plane.
SDN offers a centralized view of the network, giving an SDN Controller the ability to act as the “brain” of the network.
The SDN Controller relays on the information given by switches and routers via southbound APIs, and acts with what the applications set via northbound APIs.
One of the most well-known protocols used by SDN Controllers is OpenFlow, widely accepted even if it is not the only SDN implementation available. 
Centralized, programmable SDN environments can easily adjust to the rapidly changing business needs and can reduce costs, restrict
wasteful provisioning, as well as provide flexibility and innovation for networks.
%corretto
\subsection{OpenFlow}
  During 2006 ,thanks to M. Casado, a new network architecture was developed, called Ethane.
  On this conncept was born, during 2011, the Open Network Foundation, a no profit association that would innovate SDN and standardize OpenFlow protocol whit relatives tecnology \cite{ONF}.
  The first version of OpenFlow is 1.0.0 was released on 2009 but the last standardized version v1.5.1 was approved only on march 2015 \cite{ONF_report}.
  %corrects
  %todo spostare l'immagine correttamente
\begin{center}
 \centering
 \includegraphics[scale=0.58]{images/of.jpg}
 {Figure 1: \emph{OpenFlow.}}
 %\label{fig:topo}
\end{center}
\textbf{Openflow 1.0 messages}

@@

\subsection{Router Board}
\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.30]{images/rboard.png}

  \label{fig:topo}
\end{figure}
\begin{center}
 {Figure 2: \emph{Experience scenario.}}
\end{center}

@@

When we set up this option \texttt{OFPP_FLOOD} like output port's action in a table's switch row if some spanning_tree algorithm is working
the \texttt{OFPP_FLOOD} is only theoretical in the sense that the algorithm excludes some ports during the flooding option, in order to avoid loop\cite{pox}.
\end{itemize}
\subsection{Real Lab Scenario}
The topological scenario in which we have set up experience is as follows:
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/topo.png}
 {Figure 3: \emph{Experience scenario.}}
 \label{fig:topo}
\end{figure}

@@

  \begin{figure}[!h]
 \centering
 \includegraphics[scale=0.30]{images/topology.png}
 {Figure 4: \emph{Experience scenario.}}
 \label{fig:topo}
  \end{figure}

\subsection{Traffic's Analysis}\label{subsec:traf}
We used laptop connected via
GigabitEthernet interface and Ethernet cable cat.5a to receive the traffic from the destination host.
Our laptop has been connected via Wi-Fi to our access point via protocol 802.11n.
No other traffic sources were active during the testing, so the router was handling only
the traffic from laptop when sniffing traffic. The connection used was an ADSL2+ via copper.
We did 3 type of traffic analysis: Gaming traffic, Skype® and torrent traffic. We analyzed
300 Mb for the gaming traffic, 70 Mb for the VoIP traffic and 3GB for the torrent traffic.
The analysis involved the average number of packets per second, the packet size, the protocol (TCP or UDP).

@@

\subsubsection{Delay}
Monitoring round-trip time provides important insights for network troubleshooting and traffic engineering.
The  common  monitoring  technique  is  to  actively  send  probe packets  from  selected  vantage  points  hosts  or  middleboxes.
In traditional  networks, the control over the network routing is
limited,  making  it  impossible  to  monitor  every  selected  path. The  emerging  concept  of  Software  Defined  Networking  simplifies
network  control.  However,  OpenFlow,  the  common  SDN protocol,   does   not   support   RTT   monitoring   as   part   of   its
specification. In this experience we have tryed to emply an tecnic to escape this problem\cite{rtt}.
\newline The idea is to measure the time that occur from sending a packet from controller to a switch and the time in which the packet return from
an other switch.
\\
\newline First we have to discovery round trip time from controller to every single switch on our topology.

In order to do this on the controller we start a timer, we send a proper packet to the switch specifing the \texttt{OFPP_CONTROLLER - Send to the controller}
option and when the message return to the controller we stop the timer and calculare RTT for the switch.
\begin{center}
 \centering
 \includegraphics[scale=0.70]{images/rtt2.png}

 \label{fig:topo}

\end{center}

\begin{center}
 {Figure 5: \emph{RTT between controller and OF switch.}}
\end{center}

If we need to calculate the delay on a specific link, the rtt has to calculate with the headed switch of this link.

Now we create a proper message with unique id. number, let's start a timer \texttt{TM1} and then send this message to a first switch specifying the output port.
Then when the packet comes to the second switch it identify the particular message and sends to the controller that stops it's dedicated timer \texttt{TM2}.
\newline This part required a particular flow entry on the second switch in order to send to the controller only some particular packet.
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/rtt1.png}
 \label{fig:topo}
\end{figure}

\begin{center}
  {Figure 6: \emph{Pck from controller to itself following a particular way.}}
\end{center}

At this point we can calculate the delay of the link.
\\
\\
\begin{center}
  delay = (TM2 - TM1) - RTT1/2 - RTT2/2
\end{center}

Theoretically this method must works but in practice there are a lot of delay introduced from computational congestion on the virtualized controller.
What we observe is a estimated delay like a few seconds which is not the case, in fact if we made some \texttt{ping test} from one host to another host
in this SDN network we realize delay around 3ms.
\\
\\
The following picture shows the visual output that we can obtain from delay's graph. In this graph we storage only the core information;
\begin{center}
 \centering
 \includegraphics[scale=0.26]{images/delay.png}
 {Figure 7:\emph{ Delay's graph.}}
 \label{fig:topo}
\end{center}

%\end{figure}
\subsubsection{Link Capacity}
